# Analysis

```{r test, echo = FALSE, message = FALSE}
source(here::here("scripts/Clean/setup.R"))
```

## Answers to the research questions

### Focus on relationship between SDGs
**How are the different SDGs linked? (We want to see if some SDGs are linked in the fact that a high score on one implies a high score on the other, and thus if we can make groups of SDGs that are comparable in that way).**

## Focus on relationship between SDGs
```{r Improting the data, echo=FALSE, message=FALSE}
source(here::here("scripts","Clean","setup.R"))
threashold_heatmap <- 0.5
```

Let's analyse our relationship between the SDGs. For that, we'll import our dataset to examine the interconnections among the Sustainable Development Goals (SDGs). After importing, we'll focus specifically on the columns representing the goals of interest. To provide a comprehensive analysis, we will construct a correlation matrix, highlighting only those goals where the correlation coefficient is either greater than `r threashold_heatmap` (indicating a strong positive relationship) or less than -`r threashold_heatmap` (signifying a strong negative relationship). This approach will enable us to identify and analyze the most significant relationships between the selected SDGs.

```{r Preparation of the data, echo = TRUE, message = FALSE, warning=FALSE}
data_4 <- read.csv(here::here("scripts", "data", "data_question24.csv"))
goals_data_4_cl <- na.omit(data_4, cols=c("goal1", "goal10"))
goals_data_4_cl <- goals_data_4_cl[, grepl("goal", names(goals_data_4_cl))]
```

Given that our variables do not follow a normal distribution, employing the Pearson correlation method is not suitable in our analysis. We attempted to normalize the data through logarithmic and square root transformations, but these adjustments were insufficiently effective. Consequently, we will resort to computing the Spearman correlation. While not ideal, this method does not necessitate the normal distribution of our data. In our analysis, particularly for the heatmap visualization, we will focus on correlations that exceed the threshold of r threshold_heatmap or fall below -r threshold_heatmap. This selective approach will enhance the readability and interpretability of the heatmap.

```{r Spearman cor, echo = TRUE, message = FALSE, warning=FALSE}
spearman_corr_4_cl <- cor(goals_data_4_cl, method = "spearman", use = "everything")
spearman_corr_4_cl[abs(spearman_corr_4_cl) < threashold_heatmap] <- NA
```

We can then plot the Heatmap of the Spearman correlation using the `ggplot2` package.

```{r Spearman Heatmap, echo = TRUE, message = FALSE, warning=FALSE}
# Melting the data
melted_corr_4 <- melt(spearman_corr_4_cl, na.rm = TRUE)

# Creating the heatmap
ggplot(data = melted_corr_4, aes(x = Var1, y = Var2, fill = value)) +
    geom_tile() +
    geom_text(aes(label = sprintf("%.2f", value)), vjust = 0.5, size=2.5) + # Adding text
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                         midpoint = 0, limit = c(-1,1), space = "Lab", 
                         name="Spearman\nCorrelation",
                         na.value = "grey") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = "Heatmap of Spearman Correlations for Goals", 
         x = "", y = "")
```

It is evident that the Sustainable Development Goals (SDGs) are intricately interconnected. However, certain goals appear to be less interrelated compared to others. Specifically, SDG 1 (No Poverty) and SDG 10 (Reduced Inequalities) demonstrate a weaker correlation with the rest of the goals. Similarly, Goal 15 (Life on Land) also exhibits a lesser degree of interconnection with the other SDGs.

```{r Scree Plot and Biplot, echo = TRUE, message = FALSE, warning=FALSE}
# Selecting only numeric columns, assuming they are named as 'goal1', 'goal2', etc.
goals_data <- goals_data_4_cl[, grep('goal', names(goals_data_4_cl))]
goals_data_scaled <- scale(goals_data) # Scaling the data
pca_result <- prcomp(goals_data_scaled) # Running PCA

# Plotting Scree plot to visualize the importance of each principal component
fviz_eig(pca_result,
         addlabels = TRUE) +
  theme_minimal()

# Plotting Biplot to visualize the two main PCs
fviz_pca_biplot(pca_result,
                label="var",
                col.var="dodgerblue3",
                geom="point",
                pointsize = 0.1,
                labelsize = 5) +
  theme_minimal()
```

In our EDA on the focus on the influence of the factors over the SDG scores, we had made a correlation matrix heatmap that took into account every variable of our dataset. Here, we tried to zoom on certain parts of the heatmap. We have decided to add on our graphs the correlations between variables when our pvalue was significant (alpha = 0.05). The grey zones are concerning our non-significant pvalues. 

Let's see first the correlation matrix heatmap regarding our SDG goals and all our variables different than our SDG goals.

```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.width="100%"}
corr_matrix <- cor(data_question1[7:40])
p_matrix2 <- matrix(nrow = ncol(data_question1[7:40]), ncol = ncol(data_question1[7:40]))
for (i in 1:ncol(data_question1[7:40])) {
  for (j in 1:ncol(data_question1[7:40])) {
    test_result <- cor.test(data_question1[7:40][, i], data_question1[7:40][, j])
    p_matrix2[i, j] <- test_result$p.value
  }
}

#Switch population at the end of heatmap
corr_matrix[which(p_matrix2 > 0.05)] <- NA
melted_corr_matrix_GVar <- melt(corr_matrix[19:34,1:18])
ggplot(melted_corr_matrix_GVar, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = ifelse(!is.na(value), sprintf("%.2f", value), '')),
            color = "black", size = 2) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white",
                       midpoint = 0, limit = c(-1, 1), space = "Lab",
                       name = "Pearson\nCorrelation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_text(angle = 45, hjust = 1)) +
  labs(x = 'Goals', y = 'Goals',
       title = 'Correlations Heatmap between goals and our other variables')
```
As we can see, our SDG goals 12 & 13 (responsible consumption & production and climate action) are negatively correlated with most of our variables, as is the economic freedom government variable to our SDG goals. In that sens, we could understand it as having a higher Human Freedom Index Score would influence more negatively the SDG scores of these two goals, i.e. the more people in a country can access and afford civil justice, the more it impact negatively the score of these two SDG goals. 

Nevertheless, goals 12 & 13 and ef_government are positively correlated together. In addition, some variables such as internet_usage, pf_law or ef_legal are strongely correlated with most of our SDG goals. This is mostly due to the large scope englobed in these variables. That makes them influence various sectors of our economies and thus, mostly impacting all our SDG goals. 

Now let's zoom on the correlations between all our variables except our SDG goals:
```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.width="100%"}
melted_corr_matrix_Var <- melt(corr_matrix[19:34,19:34])
ggplot(melted_corr_matrix_Var, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = ifelse(!is.na(value), sprintf("%.2f", value), '')),
            color = "black", size = 1.7) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white",
                       midpoint = 0, limit = c(-1, 1), space = "Lab",
                       name = "Pearson\nCorrelation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_text(angle = 45, hjust = 1)) +
  labs(x = 'Goals', y = 'Goals',
       title = 'Correlations Heatmap between other variables than SDG goals')
```

```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.height="100%", out.width="100%"}
#### PCA ####

# for goals
myPCA_g <- PCA(data_question1[,9:24], graph = FALSE)
plot(myPCA_g$ind$coord[, 1], myPCA_g$ind$coord[, 2], xlab = "PC1", ylab = "PC2", main = "PCA Plot SDG Goals", pch = 19, col = "blue", cex = 0.5) + 
  abline(h = 0, col = "red", lty = 2) + 
  abline(v = 0, col = "red", lty = 2)
plot.PCA(myPCA_g, choix = "var", pch = 10, cex = 0.6)

summary(myPCA_g)
myPCA_g$eig
```

Concerning the SDG goals, we conclude that most of our variables are going along the 1st component, except the goals 10 and 15 that are rather uncorrelated with the dimension 1. In addition, as seen before, the goals 12 and 13 are negatively correlated to the other goals. With a eigenvalue bigger than 1 for the first two components, we conclude that there are only 2 dimensions to take into account, according to the Kaiser-Guttman's rule. Nevertheless, they are explaining less than 80% of cumulated variance.

```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.height="100%", out.width="100%"}
#for HFI scores
myPCA_s <- PCA(data_question1[,29:40], graph = FALSE)
plot(myPCA_s$ind$coord[, 1], myPCA_s$ind$coord[, 2], xlab = "PC1", ylab = "PC2", main = "PCA Plot HFI Scores", pch = 19, col = "blue", cex = 0.5) + 
  abline(h = 0, col = "red", lty = 2) + 
  abline(v = 0, col = "red", lty = 2)
plot.PCA(myPCA_s, choix = "var",cex = 0.5)
summary(myPCA_s)
myPCA_s$eig
```

Now concerning the Human Freedom Index scores, most of the variables are positively correlated to the dimension 1, slightly less for the PF religion and security, and finaly the EF government variable is uncorrelated to the dimension 1. With a eigenvalue bigger than 1 for the three first components, we conclude that there are 3 dimensions to take into account. Nevertheless, again, they are explaining less than 80% of cumulated variance.

```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.height="100%", out.width="100%"}
#### Kmean clustering ####

data1_scaled <- scale(Correlation_overall)
rownames(data1_scaled) <- seq_along(row.names(data1_scaled))
fviz_nbclust(data1_scaled, kmeans, method="wss")
kmean <- kmeans(data1_scaled, 7, nstart = 25)
print(kmean)
fviz_cluster(kmean, data=data1_scaled, repel=FALSE, depth =NULL, ellipse.type = "norm", labelsize = 0, pointsize = 0.5)

### NOW CLUSTERING BY COUNTRY? AND TAKE MEAN OF EVERY VARIABLE ON EVERY CONCERNED YEAR?
```

Due to the large number of data, the visualization of the clusters using the kmean method is not really relevant. In addition, by clustering our data, we are trying to get group that differ from eachother but with little variation of the observations within the same cluster. Here, only 60.6% of the variance is explained by the variation between clusters. This is not enough.

We have noticed that we had high multicolinearity in on regression. Therefore, before to compute them, let's try to get rid of on of the two variables having at least |0.8| of correlation. 
```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.width="100%"}
correlation_overall_matrix <- cor(Correlation_overall, use = "everything")
high_cor_pairs <- which(abs(correlation_overall_matrix) >= 0.8, arr.ind = TRUE)

# Displaying the results
for (i in 1:nrow(high_cor_pairs)) {
  row <- high_cor_pairs[i, "row"]
  col <- high_cor_pairs[i, "col"]
  
  # Avoiding duplicate pairs and diagonal elements
  if (row < col) {
    cat(sprintf("Variables: %s and %s, Correlation: %f\n", 
                names(Correlation_overall)[row], names(Correlation_overall)[col], correlation_overall_matrix[row, col]))
  }
}

# List of high-correlation pairs
correlation_pairs <- list(
  c("overallscore", "goal1"), c("overallscore", "goal3"), c("goal1", "goal3"),
  c("overallscore", "goal4"), c("goal1", "goal4"), c("goal3", "goal4"),
  c("overallscore", "goal6"), c("goal3", "goal6"),
  c("overallscore", "goal7"), c("goal1", "goal7"), c("goal3", "goal7"), c("goal4", "goal7"), c("goal6", "goal7"),
  c("overallscore", "goal9"), c("goal3", "goal9"),
  c("overallscore", "goal11"), c("goal3", "goal11"), c("goal4", "goal11"), c("goal6", "goal11"), c("goal7", "goal11"),
  c("goal9", "goal12"), c("goal12", "goal13"),
  c("overallscore", "goal16"), c("goal12", "goal16"),
  c("goal9", "GDPpercapita"), c("goal12", "GDPpercapita"),
  c("overallscore", "internet_usage"), c("goal9", "internet_usage"),
  c("goal12", "pf_law"), c("goal16", "pf_law"),
  c("pf_religion", "pf_assembly"), c("pf_assembly", "pf_expression"),
  c("goal9", "ef_legal"), c("goal12", "ef_legal"), c("goal16", "ef_legal"), c("pf_law", "ef_legal")
)

# Flatten the list and count the frequency of each variable
flattened_list <- unlist(correlation_pairs)
frequency_count <- table(flattened_list)
variables_to_remove <- c()

for (pair in correlation_pairs) {
  # Select the variable that appears more frequently for removal
  if (frequency_count[pair[1]] > frequency_count[pair[2]]) {
    variables_to_remove <- c(variables_to_remove, pair[1])
  } else if (frequency_count[pair[1]] < frequency_count[pair[2]]) {
    variables_to_remove <- c(variables_to_remove, pair[2])
  } else {
    # If both appear equally, arbitrarily choose one to remove
    variables_to_remove <- c(variables_to_remove, pair[1])
  }
}

variables_to_remove <- unique(variables_to_remove)
variables_to_remove <- sort(variables_to_remove)
print(variables_to_remove) 
```
Therefore, we will not take into account the variables "ef_legal"     "goal11"       "goal12"       "goal16"       "goal3"        "goal4"        "goal7"        "goal9"        "overallscore" "pf_assembly" in our regressions, for multicollinearity purpose. 

Now, let's compute the regressions without these variables. 
```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.width="100%"}
reg_goal1_all_new <- lm(goal1 ~ goal2 + goal5 + goal6 + goal8 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal2_all_new <- lm(goal2 ~ goal1 + goal5 + goal6 + goal8 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal3_all_new <- lm(goal3 ~ goal1 + goal2 + goal5 + goal6 + goal8 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal4_all_new <- lm(goal4 ~ goal1 + goal2 + goal5 + goal6 + goal8 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal5_all_new <- lm(goal5 ~ goal1 + goal2 + goal6 + goal8 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal6_all_new <- lm(goal6 ~ goal1 + goal2 + goal5 + goal8 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal7_all_new <- lm(goal7 ~ goal1 + goal2 + goal5 + goal6 + goal8 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal8_all_new <- lm(goal8 ~ goal1 + goal2 + goal5 + goal6 + goal7 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal9_all_new <- lm(goal9 ~ goal1 + goal2 + goal5 + goal6 + goal7 + goal8 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal10_all_new <- lm(goal10 ~ goal1 + goal2 + goal5 + goal6 + goal7 + goal8 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal11_all_new <- lm(goal11 ~ goal1 + goal2 + goal5 + goal6 + goal7 + goal8 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal12_all_new <- lm(goal12 ~ goal1 + goal2 + goal5 + goal6 + goal7 + goal8 + goal10 + goal11 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal13_all_new <- lm(goal13 ~ goal1 + goal2 + goal5 + goal6 + goal7 + goal8 + goal10 + goal11 + goal12 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal15_all_new <- lm(goal15 ~ goal1 + goal2 + goal5 + goal6 + goal7 + goal8 + goal10 + goal11 + goal12 + goal13 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal16_all_new <- lm(goal16 ~ goal1 + goal2 + goal5 + goal6 + goal7 + goal8 + goal10 + goal11 + goal12 + goal13 + goal15 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal17_all_new <- lm(goal17 ~ goal1 + goal2 + goal5 + goal6 + goal7 + goal8 + goal10 + goal11 + goal12 + goal13 + goal15 + goal16 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)


```
The problem is that even by getting rid of the previous variables, there still might be multicollinearity. Therefore, we need to analyse the vif for each regression and adapt the model in consequence. 
```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.width="100%"}
#for reg1
nullmod <- lm(goal1 ~ 1, data = data_question1)
selmod <- step(reg_goal1_all_new, scope=list(lower=nullmod, upper=reg_goal1_all_new), direction="backward") 
summary(selmod)
vif(selmod) #pf_law -> get rid of it
reg_goal1_all_new <- lm(goal1 ~ goal5 + goal6 + goal8 + goal10 + goal13 + goal15 + goal17 + 
                          unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + 
                          internet_usage + pf_movement + pf_religion + ef_government + 
                          ef_money + ef_trade + ef_regulation, data = data_question1)
selmod <- step(reg_goal1_all_new, scope=list(lower=nullmod, upper=reg_goal1_all_new), direction="backward") 
vif(selmod)
reg_goal1_all_new <- lm(goal1 ~ goal5 + goal6 + goal8 + goal10 + goal13 + goal15 + goal17 + 
                          unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + 
                          internet_usage + pf_movement + pf_religion + ef_government + 
                          ef_money + ef_trade + ef_regulation, data = data_question1)
#for reg2
nullmod <- lm(goal1 ~ 1, data = data_question1)
selmod <- step(reg_goal2_all_new, scope=list(lower=nullmod, upper=reg_goal2_all_new), direction="backward") 
vif(selmod) 
reg_goal2_all_new <- lm(goal2 ~ goal5 + goal6 + goal8 + goal10 + goal13 + goal15 + goal17 + 
                          unemployment.rate + MilitaryExpenditurePercentGDP + internet_usage + 
                          pf_law + pf_security + pf_movement + pf_identity + ef_money + 
                          ef_trade + ef_regulation + population, data = data_question1)
#reg5
nullmod <- lm(goal1 ~ 1, data = data_question1)
selmod <- step(reg_goal5_all_new, scope=list(lower=nullmod, upper=reg_goal5_all_new), direction="backward") 
vif(selmod) #goal6
reg_goal5_all_new <- lm(goal5 ~ goal1 + goal2 + goal10 + goal13 + goal15 + goal17 + 
                          unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + 
                          internet_usage + pf_law + pf_security + pf_movement + pf_religion + 
                          pf_expression + pf_identity + ef_government + ef_money + 
                          ef_trade + ef_regulation, data = data_question1)
selmod <- step(reg_goal5_all_new, scope=list(lower=nullmod, upper=reg_goal5_all_new), direction="backward") 
vif(selmod) #pf_law
reg_goal5_all_new <- lm(goal5 ~ goal1 + goal2 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + 
                          GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + 
                          pf_security + pf_movement + pf_religion + pf_identity + 
                          ef_government + ef_money + ef_trade + ef_regulation, data = data_question1)
selmod <- step(reg_goal5_all_new, scope=list(lower=nullmod, upper=reg_goal5_all_new), direction="backward") 
vif(selmod) 
reg_goal5_all_new <- lm(goal5 ~ goal1 + goal2 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + 
                          GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + 
                          pf_security + pf_movement + pf_religion + pf_identity + ef_government + 
                          ef_money + ef_regulation, data = data_question1)
#reg6
nullmod <- lm(goal1 ~ 1, data = data_question1)
selmod <- step(reg_goal6_all_new, scope=list(lower=nullmod, upper=reg_goal6_all_new), direction="backward") 
vif(selmod) 
reg_goal6_all_new <- lm(goal6 ~ goal1 + goal2 + goal5 + goal8 + unemployment.rate + GDPpercapita + 
                          MilitaryExpenditurePercentGDP + internet_usage + pf_security + 
                          pf_movement + pf_religion + pf_expression + pf_identity + 
                          ef_government + ef_money + population, data = data_question1)

#reg8
nullmod <- lm(goal1 ~ 1, data = data_question1)
selmod <- step(reg_goal8_all_new, scope=list(lower=nullmod, upper=reg_goal8_all_new), direction="backward") 
vif(selmod) #goal6
reg_goal8_all_new <- lm(goal8 ~ goal1 + goal2 + goal13 + goal15 + unemployment.rate + 
                          internet_usage + pf_law + pf_security + pf_movement + pf_religion + 
                          pf_expression + pf_identity + ef_government + ef_trade + 
                          population, data = data_question1)

#reg10
nullmod <- lm(goal1 ~ 1, data = data_question1)
selmod <- step(reg_goal10_all_new, scope=list(lower=nullmod, upper=reg_goal10_all_new), direction="backward") 
vif(selmod) #goal6
reg_goal10_all_new <- lm(goal10 ~ goal1 + goal2 + goal5 + goal8 + goal13 + goal15 + 
                           goal17 + unemployment.rate + GDPpercapita + internet_usage + 
                           pf_law + pf_security + pf_movement + pf_religion + pf_expression + 
                           ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
selmod <- step(reg_goal10_all_new, scope=list(lower=nullmod, upper=reg_goal10_all_new), direction="backward") 
vif(selmod) #pf_law
reg_goal10_all_new <- lm(goal10 ~ goal1 + goal2 + goal5 + goal8 + goal13 + goal15 + goal17 + 
                           unemployment.rate + GDPpercapita + internet_usage + 
                           pf_security + pf_movement + pf_religion + pf_expression + 
                           ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)

#reg13
nullmod <- lm(goal1 ~ 1, data = data_question1)
selmod <- step(reg_goal13_all_new, scope=list(lower=nullmod, upper=reg_goal13_all_new), direction="backward") 
vif(selmod) 
reg_goal13_all_new <- lm(goal13 ~ goal1 + goal2 + goal5 + goal8 + goal10 + goal17 + unemployment.rate + 
                           GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + 
                           pf_law + pf_religion + ef_government + ef_regulation, data = data_question1)

#reg15
nullmod <- lm(goal1 ~ 1, data = data_question1)
selmod <- step(reg_goal15_all_new, scope=list(lower=nullmod, upper=reg_goal15_all_new), direction="backward") 
vif(selmod) 
reg_goal15_all_new <- lm(goal15 ~ goal1 + goal2 + goal5 + goal8 + goal10 + goal17 + unemployment.rate + 
                           GDPpercapita + internet_usage + pf_law + pf_security + pf_religion + 
                           pf_expression + pf_identity + ef_government + population, data = data_question1)

#reg16
nullmod <- lm(goal1 ~ 1, data = data_question1)
selmod <- step(reg_goal16_all_new, scope=list(lower=nullmod, upper=reg_goal16_all_new), direction="backward") 
vif(selmod) 
reg_goal16_all_new <- lm(goal16 ~ goal1 + goal2 + goal5 + goal8 + goal10 + goal13 + 
                           unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + 
                           internet_usage + pf_law + pf_security + pf_movement + pf_religion + 
                           pf_expression + pf_identity + ef_government + ef_money + 
                           ef_regulation + population, data = data_question1)
selmod <- step(reg_goal16_all_new, scope=list(lower=nullmod, upper=reg_goal16_all_new), direction="backward") 
vif(selmod)#pf_law
reg_goal16_all_new <- lm(goal16 ~ goal1 + goal2 + goal5 + goal8 + goal10 + goal13 + unemployment.rate + 
  GDPpercapita + MilitaryExpenditurePercentGDP + pf_security + 
  pf_movement + pf_religion + pf_expression + pf_identity + 
  ef_government + ef_money + ef_regulation + population, data = data_question1)
selmod <- step(reg_goal16_all_new, scope=list(lower=nullmod, upper=reg_goal16_all_new), direction="backward") 
vif(selmod) #pf_law
reg_goal16_all_new <- lm(goal16 ~ goal1 + goal2 + goal5 + goal8 + goal10 + goal13 + unemployment.rate + 
                           GDPpercapita + MilitaryExpenditurePercentGDP + pf_security + 
                           pf_movement + pf_religion + pf_expression + pf_identity + 
                           ef_government + ef_money + ef_regulation + population, data = data_question1)

#reg17
nullmod <- lm(goal1 ~ 1, data = data_question1)
selmod <- step(reg_goal17_all_new, scope=list(lower=nullmod, upper=reg_goal17_all_new), direction="backward") 
vif(selmod) 
reg_goal17_all_new <- lm(goal17 ~ goal1 + goal2 + goal5 + goal10 + goal13 + goal15 + unemployment.rate + 
                           GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_security + pf_movement + pf_religion + pf_expression +
                           pf_identity + ef_government + ef_money + ef_trade + ef_regulation + 
                           population, data=data_question1)
```
stargazer regressions
```{r results='asis'}
sg1 <- stargazer(reg_goal1_all_new,
                 reg_goal2_all_new,
          title="Impact of variables over SDG goals 1,2",
          type='html',
          digits=3)
```

```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.width="100%"}
##### geom point #####

#print values with correlation > 0.8 and make plots

# Filtering values where the absolute value is greater than 0.8
highcorrelations <- melted_corr_matrix_GVar %>% filter(value > 0.8)

ggplot(data_question1, aes(internet_usage, overallscore)) +
  geom_point()+ geom_smooth(se = FALSE) +
  labs(title = "Scarplot overallscore and internet usage")

ggplot(data_question1, aes(GDPpercapita, goal9)) +
  geom_point()+ geom_smooth(se = FALSE) +
  labs(title = "Scarplot overallscore and internet usage")

ggplot(data_question1, aes(internet_usage,goal9)) +
  geom_point()+ geom_smooth(se = FALSE) +
  labs(title = "Scarplot overallscore and internet usage")

ggplot(data_question1, aes(ef_legal,goal9)) +
  geom_point()+ geom_smooth(se = FALSE) +
  labs(title = "Scarplot overallscore and internet usage")

ggplot(data_question1, aes(pf_law, goal16)) +
  geom_point()+ geom_smooth(se = FALSE) +
  labs(title = "Scarplot overallscore and internet usage")

ggplot(data_question1, aes(ef_legal, goal16)) +
  geom_point()+ geom_smooth(se = FALSE) +
  labs(title = "Scarplot overallscore and internet usage")
```

Let's explore how the different SDG are correlated together by creating a heatmap of the correlation between our variables. We added a script to check whether the correlations are significantly different from 0. First, let's select the SDGs scores.

```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.width="100%"}
sdg_scores <- Q4[, c('goal1', 'goal2', 'goal3', 'goal4', 'goal5', 'goal6',
                     'goal7', 'goal8', 'goal9', 'goal10', 'goal11', 'goal12',
                     'goal13', 'goal15', 'goal16', 'goal17')]
```

We then, initialize the matrices and calculate the correlation, and p-values of each combination of SDGs scores

```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.width="100%"}
cor_matrix <- matrix(nrow = ncol(sdg_scores), ncol = ncol(sdg_scores))
p_matrix <- matrix(nrow = ncol(sdg_scores), ncol = ncol(sdg_scores))
rownames(cor_matrix) <- colnames(sdg_scores)
rownames(p_matrix) <- colnames(sdg_scores)
colnames(cor_matrix) <- colnames(sdg_scores)
colnames(p_matrix) <- colnames(sdg_scores)

# Calculate correlation and p-values
for (i in 1:ncol(sdg_scores)) {
  for (j in 1:ncol(sdg_scores)) {
    test_result <- cor.test(sdg_scores[, i], sdg_scores[, j])
    cor_matrix[i, j] <- test_result$estimate
    p_matrix[i, j] <- test_result$p.value}}
```

We then reshape our data to be able to use the `ggplot2` package to create our heatmap.

```{r, echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, out.width="100%"}
melted_cor_matrix <-
  melt(cor_matrix)
melted_p_matrix <-
  melt(matrix(as.vector(p_matrix), nrow = ncol(sdg_scores)))

plot_data <- # Combine the datasets
  cbind(melted_cor_matrix, p_value = melted_p_matrix$value)

ggplot(plot_data, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f", value), color = p_value < 0.05),
            vjust = 1) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
  scale_color_manual(values = c("black", "yellow")) + # black when significant, yellow if not
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_text(angle = 45, hjust = 1),
        legend.position = "none") +
  labs(x = 'SDG Goals', y = 'SDG Goals',
       title = 'Correlation Matrix with Significance Indicator')
```

Note that as said previously, we assessed the correlations to ascertain if they substantially deviated from zero, setting the significance level at an alpha of 5%. To aid in visualization, we marked any correlations that did not meet this level of significance with a yellow highlight in our graphical representation. The absence of yellow markings on our plot suggests that all Sustainable Development Goal (SDG) scores demonstrate a statistically significant correlation.

We can have a look at the shape of the corelation between the SDGs with the plot function.

```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.width="100%"}
plot(sdg_scores)
```



## Different methods considered

## Competing approaches

## Justifications
