# Analysis

```{r test, echo = FALSE, message = FALSE}
source(here::here("scripts/Clean/setup.R"))
```

## Answers to the research questions

### Focus on relationship between SDGs
**How are the different SDGs linked? (We want to see if some SDGs are linked in the fact that a high score on one implies a high score on the other, and thus if we can make groups of SDGs that are comparable in that way).**

In our EDA on the focus on the influence of the factors over the SDG scores, we had made a correlation matrix heatmap that took into account every variable of our dataset. Here, we tried to zoom on certain parts of the heatmap. We have decided to add on our graphs the correlations between variables when our pvalue was significant (alpha = 0.05). The grey zones are concerning our non-significant pvalues. 

Let's see first the correlation matrix heatmap regarding our SDG goals and all our variables different than our SDG goals.

```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.width="100%"}
corr_matrix <- cor(data_question1[7:40])
p_matrix2 <- matrix(nrow = ncol(data_question1[7:40]), ncol = ncol(data_question1[7:40]))
for (i in 1:ncol(data_question1[7:40])) {
  for (j in 1:ncol(data_question1[7:40])) {
    test_result <- cor.test(data_question1[7:40][, i], data_question1[7:40][, j])
    p_matrix2[i, j] <- test_result$p.value
  }
}

#Switch population at the end of heatmap

corr_matrix[which(p_matrix2 > 0.05)] <- NA
melted_corr_matrix_GVar <- melt(corr_matrix[19:34,1:18])
ggplot(melted_corr_matrix_GVar, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = ifelse(!is.na(value), sprintf("%.2f", value), '')),
            color = "black", size = 2) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white",
                       midpoint = 0, limit = c(-1, 1), space = "Lab",
                       name = "Pearson\nCorrelation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_text(angle = 45, hjust = 1)) +
  labs(x = 'Goals', y = 'Goals',
       title = 'Correlations Heatmap between goals and our other variables')
```
As we can see, our SDG goals 12 & 13 (responsible consumption & production and climate action) are negatively correlated with most of our variables, as is the economic freedom government variable to our SDG goals. In that sens, we could understand it as having a higher Human Freedom Index Score would influence more negatively the SDG scores of these two goals, i.e. the more people in a country can access and afford civil justice, the more it impact negatively the score of these two SDG goals. 

Nevertheless, goals 12 & 13 and ef_government are positively correlated together. In addition, some variables such as internet_usage, pf_law or ef_legal are strongely correlated with most of our SDG goals. This is mostly due to the large scope englobed in these variables. That makes them influence various sectors of our economies and thus, mostly impacting all our SDG goals. 

Now let's zoom on the correlations between all our variables except our SDG goals:
```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.width="100%"}
melted_corr_matrix_Var <- melt(corr_matrix[19:34,19:34])
ggplot(melted_corr_matrix_Var, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = ifelse(!is.na(value), sprintf("%.2f", value), '')),
            color = "black", size = 1.7) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white",
                       midpoint = 0, limit = c(-1, 1), space = "Lab",
                       name = "Pearson\nCorrelation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_text(angle = 45, hjust = 1)) +
  labs(x = 'Goals', y = 'Goals',
       title = 'Correlations Heatmap between other variables than SDG goals')
```
We have noticed that we had high multicolinearity in on regression. Therefore, before to compute them, let's try to get rid of on of the two variables having at least |0.8| of correlation. 
```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.width="100%"}
correlation_overall_matrix <- cor(Correlation_overall, use = "everything")
high_cor_pairs <- which(abs(correlation_overall_matrix) >= 0.8, arr.ind = TRUE)

# Displaying the results
for (i in 1:nrow(high_cor_pairs)) {
  row <- high_cor_pairs[i, "row"]
  col <- high_cor_pairs[i, "col"]
  
  # Avoiding duplicate pairs and diagonal elements
  if (row < col) {
    cat(sprintf("Variables: %s and %s, Correlation: %f\n", 
                names(Correlation_overall)[row], names(Correlation_overall)[col], correlation_overall_matrix[row, col]))
  }
}

# List of high-correlation pairs
correlation_pairs <- list(
  c("overallscore", "goal1"), c("overallscore", "goal3"), c("goal1", "goal3"),
  c("overallscore", "goal4"), c("goal1", "goal4"), c("goal3", "goal4"),
  c("overallscore", "goal6"), c("goal3", "goal6"),
  c("overallscore", "goal7"), c("goal1", "goal7"), c("goal3", "goal7"), c("goal4", "goal7"), c("goal6", "goal7"),
  c("overallscore", "goal9"), c("goal3", "goal9"),
  c("overallscore", "goal11"), c("goal3", "goal11"), c("goal4", "goal11"), c("goal6", "goal11"), c("goal7", "goal11"),
  c("goal9", "goal12"), c("goal12", "goal13"),
  c("overallscore", "goal16"), c("goal12", "goal16"),
  c("goal9", "GDPpercapita"), c("goal12", "GDPpercapita"),
  c("overallscore", "internet_usage"), c("goal9", "internet_usage"),
  c("goal12", "pf_law"), c("goal16", "pf_law"),
  c("pf_religion", "pf_assembly"), c("pf_assembly", "pf_expression"),
  c("goal9", "ef_legal"), c("goal12", "ef_legal"), c("goal16", "ef_legal"), c("pf_law", "ef_legal")
)

# Flatten the list and count the frequency of each variable
flattened_list <- unlist(correlation_pairs)
frequency_count <- table(flattened_list)
variables_to_remove <- c()

for (pair in correlation_pairs) {
  # Select the variable that appears more frequently for removal
  if (frequency_count[pair[1]] > frequency_count[pair[2]]) {
    variables_to_remove <- c(variables_to_remove, pair[1])
  } else if (frequency_count[pair[1]] < frequency_count[pair[2]]) {
    variables_to_remove <- c(variables_to_remove, pair[2])
  } else {
    # If both appear equally, arbitrarily choose one to remove
    variables_to_remove <- c(variables_to_remove, pair[1])
  }
}

variables_to_remove <- unique(variables_to_remove)
variables_to_remove <- sort(variables_to_remove)
print(variables_to_remove) 
```
Therefore, we will not take into account the variables "ef_legal"     "goal11"       "goal12"       "goal16"       "goal3"        "goal4"        "goal7"        "goal9"        "overallscore" "pf_assembly" in our regressions, for multicollinearity purpose. 

Now, let's compute the regressions without these variables. 
```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.width="100%"}
reg_goal1_all_new <- lm(goal1 ~ goal2 + goal5 + goal6 + goal8 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal2_all_new <- lm(goal2 ~ goal1 + goal5 + goal6 + goal8 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal3_all_new <- lm(goal3 ~ goal1 + goal2 + goal5 + goal6 + goal8 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal4_all_new <- lm(goal4 ~ goal1 + goal2 + goal5 + goal6 + goal8 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal5_all_new <- lm(goal5 ~ goal1 + goal2 + goal6 + goal8 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal6_all_new <- lm(goal6 ~ goal1 + goal2 + goal5 + goal8 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal7_all_new <- lm(goal7 ~ goal1 + goal2 + goal5 + goal6 + goal8 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal8_all_new <- lm(goal8 ~ goal1 + goal2 + goal5 + goal6 + goal7 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal9_all_new <- lm(goal9 ~ goal1 + goal2 + goal5 + goal6 + goal7 + goal8 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal10_all_new <- lm(goal10 ~ goal1 + goal2 + goal5 + goal6 + goal7 + goal8 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal11_all_new <- lm(goal11 ~ goal1 + goal2 + goal5 + goal6 + goal7 + goal8 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal12_all_new <- lm(goal12 ~ goal1 + goal2 + goal5 + goal6 + goal7 + goal8 + goal10 + goal11 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal13_all_new <- lm(goal13 ~ goal1 + goal2 + goal5 + goal6 + goal7 + goal8 + goal10 + goal11 + goal12 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal15_all_new <- lm(goal15 ~ goal1 + goal2 + goal5 + goal6 + goal7 + goal8 + goal10 + goal11 + goal12 + goal13 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal16_all_new <- lm(goal16 ~ goal1 + goal2 + goal5 + goal6 + goal7 + goal8 + goal10 + goal11 + goal12 + goal13 + goal15 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)
reg_goal17_all_new <- lm(goal17 ~ goal1 + goal2 + goal5 + goal6 + goal7 + goal8 + goal10 + goal11 + goal12 + goal13 + goal15 + goal16 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_law + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)


```
The problem is that even by getting rid of the previous variables, there still might be multicollinearity. Therefore, we need to analyse the vif for each regression and adapt the model in consequence. 
```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.width="100%"}
#first one example

nullmod <- lm(goal1 ~ 1, data = data_question1)
selmod <- step(reg_goal1_all_new, scope=list(lower=nullmod, upper=reg_goal1_all_new), direction="backward") 
vif(selmod) #only vif > 5 for pf_law -> get rid of it

#new model for goal1
reg_goal1_all_new <- lm(goal1 ~ goal2 + goal5 + goal6 + goal8 + goal10 + goal13 + goal15 + goal17 + unemployment.rate + GDPpercapita + MilitaryExpenditurePercentGDP + internet_usage + pf_security + pf_movement + pf_religion + pf_expression + pf_identity + ef_government + ef_money + ef_trade + ef_regulation + population, data = data_question1)

#and so on for goals remaining after deleting variables when correlations > |0.8| (goal2 + goal5 + goal6 + goal8 + goal10 + goal13 + goal15 + goal17)


```
stargazer reg 1
```{r results='asis'}
sg1 <- stargazer(reg_goal1_all_new,
          title="Impact of variables over SDG goals1",
          type='html',
          digits=3)
```
Let's explore how the different SDG are correlated together by creating a heatmap of the correlation between our variables. We added a script to check whether the correlations are significantly different from 0. First, let's select the SDGs scores.

```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.width="100%"}
sdg_scores <- Q4[, c('goal1', 'goal2', 'goal3', 'goal4', 'goal5', 'goal6',
                     'goal7', 'goal8', 'goal9', 'goal10', 'goal11', 'goal12',
                     'goal13', 'goal15', 'goal16', 'goal17')]
```

We then, initialize the matrices and calculate the correlation, and p-values of each combination of SDGs scores

```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.width="100%"}
cor_matrix <- matrix(nrow = ncol(sdg_scores), ncol = ncol(sdg_scores))
p_matrix <- matrix(nrow = ncol(sdg_scores), ncol = ncol(sdg_scores))
rownames(cor_matrix) <- colnames(sdg_scores)
rownames(p_matrix) <- colnames(sdg_scores)
colnames(cor_matrix) <- colnames(sdg_scores)
colnames(p_matrix) <- colnames(sdg_scores)

# Calculate correlation and p-values
for (i in 1:ncol(sdg_scores)) {
  for (j in 1:ncol(sdg_scores)) {
    test_result <- cor.test(sdg_scores[, i], sdg_scores[, j])
    cor_matrix[i, j] <- test_result$estimate
    p_matrix[i, j] <- test_result$p.value}}
```

We then reshape our data to be able to use the `ggplot2` package to create our heatmap.

```{r, echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, out.width="100%"}
melted_cor_matrix <-
  melt(cor_matrix)
melted_p_matrix <-
  melt(matrix(as.vector(p_matrix), nrow = ncol(sdg_scores)))

plot_data <- # Combine the datasets
  cbind(melted_cor_matrix, p_value = melted_p_matrix$value)

ggplot(plot_data, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f", value), color = p_value < 0.05),
            vjust = 1) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
  scale_color_manual(values = c("black", "yellow")) + # black when significant, yellow if not
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_text(angle = 45, hjust = 1),
        legend.position = "none") +
  labs(x = 'SDG Goals', y = 'SDG Goals',
       title = 'Correlation Matrix with Significance Indicator')
```

Note that as said previously, we assessed the correlations to ascertain if they substantially deviated from zero, setting the significance level at an alpha of 5%. To aid in visualization, we marked any correlations that did not meet this level of significance with a yellow highlight in our graphical representation. The absence of yellow markings on our plot suggests that all Sustainable Development Goal (SDG) scores demonstrate a statistically significant correlation.

We can have a look at the shape of the corelation between the SDGs with the plot function.

```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.width="100%"}
plot(sdg_scores)
```



## Different methods considered

## Competing approaches

## Justifications
