# Exploratory data analysis

```{r setup, echo = FALSE, message = FALSE, warning=FALSE}
source(here::here("scripts/Clean/setup.R"))
source(here::here("scripts/EDA/EDA_4_SDG.R"))
data_question1 <- read.csv(here("scripts","data","data_question1.csv"))
data_question24 <- read.csv(here("scripts", "data", "data_question24.csv"))
data_question2 <- read.csv(here("scripts", "data", "data_question24.csv"))
data_question3_1 <- read.csv(here("scripts", "data", "data_question3_1.csv"))
data_question3_2 <- read.csv(here("scripts", "data", "data_question3_2.csv"))
data_question3_3 <- read.csv(here("scripts", "data", "data_question3_3.csv"))
Q3.1 <- read.csv(here("scripts", "data", "data_question3_1.csv"))
Q3.2 <- read.csv(here("scripts", "data", "data_question3_2.csv"))
Q3.3 <- read.csv(here("scripts", "data", "data_question3_3.csv"))
data <- read.csv(here("scripts", "data", "all_Merge.csv"))
Correlation_overall <- data_question1 %>% 
      dplyr::select(population:ef_regulation)
```

## Focus on the influence of the factors over the SDG scores

For this first part of our EDA, let's try to check first the distribution of the variables selected for answering our 1st question. 

```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.height="100%", out.width="100%"}

# Reshape the data from wide to long format for our sdg goals and our human freedom index scores
long_df_goal_distribution <- pivot_longer(Correlation_overall, cols = starts_with("goal"), names_to = "Goal", values_to = "Value")

long_df_hfi_distribution <- pivot_longer(Correlation_overall, cols = pf_law:ef_regulation, names_to = "Category", values_to = "Value")

ggplot(long_df_goal_distribution, aes(x = Value, y = Goal, fill = stat(x))) +
  geom_density_ridges_gradient(scale = 3, size = 0.3, rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "", option = "C") +
  theme(plot.title = element_text(hjust = 0.5), # Center the title
        plot.title.position = "plot") + 
  labs(x = 'Scores',
       y = 'Goals',
        title = 'SDG Goals Distribution')
```

As we can see, most of our goals have a left-skewed distribution, which shows that for most of the country concerned implemented good strategies for targeting the goals objectives. Some distribution have a longer distribution than other, which could be a proof of inequality in the investments made for implementing solutions. In another hand, we notice that the only right-skewed distribution is concerning the observations of goal 9, which is promoting infrastructures, innovation and inclusive and sustainable industrialization. Again, that could show means inequalities. Wealthier countries are able to invest more on these sustainable development goals. 

```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.height="100%", out.width="100%"}

ggplot(long_df_hfi_distribution, aes(x = Value, y = Category, fill = stat(x))) +
  geom_density_ridges_gradient(scale = 3, size = 0.3, rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "", option = "C") +
  theme(plot.title = element_text(hjust = 0.5), # Center the title
        plot.title.position = "plot") + 
  labs(x = 'Scores',
    title = 'Human Freedom Index Scores Distribution')
```

The distribution of the Human Freedom Index Score follows the same trend. Most of the scores are left-skewed, which means that countries tend to have in general good scores. The only scores not folowing are pf_law and ef_legal, which tend to have lower scores in general. Legal system, for civilians and countries, is changing slowly because it has a lot of implications over the situation within a country/between countries and because of the divergence of opinions. Therefore, investing more money for raising these scores will take more time than raising the scores of other goals. 

Now let's consider the remaining variables of the dataset dedicated to answering the influence of factors over our SDG goals scores. All these variables have right-skewed distribution. Taking the mode into account, most of the concerned countries in our data have an unemployment rate between 2 to 7%, a distribution of GDP per capita between $3'000-$10'000, a distribution of military expenditure in percentage of the GDP 10% to 20% and finally a internet usage between 0 and 10%.

These variables shows us even more the inequalities between the countries in our dataset. While most of our contries have low internet usage or/and a low GDP per capita, few countries are more developed, then mostly wealthier, and thus having better chances to get higher scores. 

```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.height="100%", out.width="100%"}
#now, same for the remaining variables. No need to reshape our data as only one variable.
unempl_d <- ggplot(Correlation_overall, aes(x = unemployment.rate, y = 1, fill = stat(x))) +
  geom_density_ridges_gradient(scale = 3, size = 0.3, rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "", option = "C") +
  theme(plot.title = element_text(hjust = 0.5, size = 10), # Center the title
        plot.title.position = "plot") + 
  labs(y = 'Density',
  title = 'Distribution of Unemployment Rate')

gdp_d <- ggplot(Correlation_overall, aes(x = GDPpercapita, y = 1, fill = stat(x))) +
  geom_density_ridges_gradient(scale = 3, size = 0.3, rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "", option = "C") +
  theme(plot.title = element_text(hjust = 0.5, size = 10), # Center the title
        plot.title.position = "plot") + 
  labs(y = 'Density', title = 'Distribution of GDP per Capita')

milit_d <- ggplot(Correlation_overall, aes(x = MilitaryExpenditurePercentGDP, y = 1, fill = stat(x))) +
  geom_density_ridges_gradient(scale = 3, size = 0.3, rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "", option = "C") +
  theme(plot.title = element_text(hjust = 0.5, size = 10), # Center the title
        plot.title.position = "plot") + 
  labs(y = 'Density',title = 'Distribution of Military Expenditure (% of GDP)')

internet_d <- ggplot(Correlation_overall, aes(x = internet_usage, y = 1, fill = stat(x))) +
  geom_density_ridges_gradient(scale = 3, size = 0.3, rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "", option = "C") +theme(plot.title = element_text(hjust = 0.5, size = 10),
        plot.title.position = "plot") + 
  labs(y = 'Density',title = 'Distribution of Internet Usage')

grid.arrange(unempl_d,gdp_d,milit_d,internet_d, ncol = 2, nrow = 2)
```
Now, let's display the distribution of the different SDG achievement scores per continent, using violin boxplots to have an overview of the mods, the range with most of the observations and the outliers.

```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.height="100%", out.width="100%"}
#### boxplots ####

#For sdg goals per continent 

#Africa
data_Q1_Africa <- data_question1 %>% #filtering Africa as continent
  filter(data_question1$continent == 'Africa') %>%
  dplyr::select(continent, overallscore, goal1, goal2, goal3, goal4, goal5, goal6, goal7, goal8, goal9, goal10, goal11, goal12, goal13, goal15, goal16, goal17)

data_Q1_Africa_long <- melt(data_Q1_Africa)

medians_AF <- data_Q1_Africa_long %>% #median per variable
  group_by(variable) %>%
  summarize(median_value = median(value))
data_Q1_Africa_long <- data_Q1_Africa_long %>%
                  left_join(medians_AF, by = "variable")
#America
data_Q1_Americas <- data_question1 %>%#filtering Americas as continent
  filter(data_question1$continent == 'Americas') %>%
  dplyr::select(continent, overallscore, goal1, goal2, goal3, goal4, goal5, goal6, goal7, goal8, goal9, goal10, goal11, goal12, goal13, goal15, goal16, goal17)

data_Q1_Americas_long <- melt(data_Q1_Americas)

medians_AM <- data_Q1_Americas_long %>% #median per variable
  group_by(variable) %>%
  summarize(median_value = median(value))
data_Q1_Americas_long <- data_Q1_Americas_long %>%
                  left_join(medians_AM, by = "variable")
#Asia
data_Q1_Asia <- data_question1 %>%
  filter(data_question1$continent == 'Asia') %>%#filtering Asia as continent
  dplyr::select(continent, overallscore, goal1, goal2, goal3, goal4, goal5, goal6, goal7, goal8, goal9, goal10, goal11, goal12, goal13, goal15, goal16, goal17)

data_Q1_Asia_long <- melt(data_Q1_Asia)

medians_AS <- data_Q1_Asia_long %>% #median per variable
  group_by(variable) %>%
  summarize(median_value = median(value))
data_Q1_Asia_long <- data_Q1_Asia_long %>%
                  left_join(medians_AS, by = "variable")
#Europe
data_Q1_Europe <- data_question1 %>%
  filter(data_question1$continent == 'Europe') %>% #filtering Europe as continent
  dplyr::select(continent, overallscore, goal1, goal2, goal3, goal4, goal5, goal6, goal7, goal8, goal9, goal10, goal11, goal12, goal13, goal15, goal16, goal17)

data_Q1_Europe_long <- melt(data_Q1_Europe)

medians_EU <- data_Q1_Europe_long %>% #median per variable
  group_by(variable) %>%
  summarize(median_value = median(value))
data_Q1_Europe_long <- data_Q1_Europe_long %>%
                  left_join(medians_EU, by = "variable")
#Oceania
data_Q1_Oceania <- data_question1 %>%
  filter(data_question1$continent == 'Oceania') %>% #filtering Oceania as continent
  dplyr::select(continent, overallscore, goal1, goal2, goal3, goal4, goal5, goal6, goal7, goal8, goal9, goal10, goal11, goal12, goal13, goal15, goal16, goal17)

data_Q1_Oceania_long <- melt(data_Q1_Oceania)

medians_OC <- data_Q1_Oceania_long %>% #median per variable
  group_by(variable) %>%
  summarize(median_value = median(value))
data_Q1_Oceania_long <- data_Q1_Oceania_long %>%
                  left_join(medians_OC, by = "variable")
# merge all medians
medians_all <- rbind(data_Q1_Oceania_long, data_Q1_Americas_long,data_Q1_Africa_long,data_Q1_Asia_long,data_Q1_Europe_long)

medians_all$color <- ifelse(medians_all$median_value > 75, "lightgreen",
                        ifelse(medians_all$median_value < 25, "red3", 'lightblue3')) #assigning colors. If median for a goal is > 75 -> lightblue, if < 25 -> red, orange otherwise.

bandwidth_nrd <- bw.nrd(medians_all$value) #adapting the bandwidth

ggplot(medians_all, aes(x = variable, y = value, fill = color)) +
  geom_violin(trim = FALSE, bw = bandwidth_nrd) +
  #geom_boxplot(width = 0.1, outlier.size = 1, fill = 'white') +
  scale_fill_identity() +
  labs(title = "SDG Goals by Region", x = "Goals", y = "Score") +
  facet_grid(continent ~ ., scales = "free_y") +
  scale_y_continuous(labels = scales::label_number()) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
Here is the distribution of the goals per continent. We notice that Europe the continent with most of its goals having a median superior to 75 (represented by the lightblue color. We notice that only two goals have a median score lower than 25, which is for goal 9 for Africa and goal 10 for America. As seen before, goal 9 is generally having lower scores than the other goals. That could mean that the access to technology and sustainable/resilient infrastructures/industrialization is harder in Africa, because of various reasons such as less wealthy countries, corruption,... 

The goal 10 concerns the reduction of inequalities within/amongst countries. Therefore, we presume that less effort and investissement has been made on this goal in America. 

In addition, some distributions are quite disparsed, such as goal 13 in Oceania and goal 10 in Africa. That could again show inequalities within countries or less investment made to raise the scores by different countriews of the same continent. 

Now let's display boxplots for the different variables of the human freedom index.

```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.height="100%", out.width="100%"}
#for Human Freedom Index scores 

#Africa
data_Q1_Africa_HFI <- data_question1 %>%
  filter(data_question1$continent == 'Africa') %>%
  dplyr::select(continent, pf_law, pf_security, pf_movement, pf_religion, pf_assembly, pf_expression, pf_identity, ef_government, ef_legal, ef_money, ef_trade, ef_regulation)

data_Q1_Africa_HFI_long <- melt(data_Q1_Africa_HFI)

medians_AF_HFI <- data_Q1_Africa_HFI_long %>%
  group_by(variable) %>%
  summarize(median_value = median(value))

data_Q1_Africa_HFI_long <- data_Q1_Africa_HFI_long %>%
                  left_join(medians_AF_HFI, by = "variable")

#America
data_Q1_Americas_HFI <- data_question1 %>%
  filter(data_question1$continent == 'Americas') %>%
  dplyr::select(continent, pf_law, pf_security, pf_movement, pf_religion, pf_assembly, pf_expression, pf_identity, ef_government, ef_legal, ef_money, ef_trade, ef_regulation)

data_Q1_Americas_HFI_long <- melt(data_Q1_Americas_HFI)

medians_AM_HFI <- data_Q1_Americas_HFI_long %>%
  group_by(variable) %>%
  summarize(median_value = median(value))

data_Q1_Americas_HFI_long <- data_Q1_Americas_HFI_long %>%
                  left_join(medians_AM_HFI, by = "variable")

#Asia
data_Q1_Asia_HFI <- data_question1 %>%
  filter(data_question1$continent == 'Asia') %>%
  dplyr::select(continent, pf_law, pf_security, pf_movement, pf_religion, pf_assembly, pf_expression, pf_identity, ef_government, ef_legal, ef_money, ef_trade, ef_regulation)

data_Q1_Asia_HFI_long <- melt(data_Q1_Asia_HFI)

medians_AS_HFI <- data_Q1_Asia_HFI_long %>%
  group_by(variable) %>%
  summarize(median_value = median(value))

data_Q1_Asia_HFI_long <- data_Q1_Asia_HFI_long %>%
                  left_join(medians_AS_HFI, by = "variable")

#Europe
data_Q1_Europe_HFI <- data_question1 %>%
  filter(data_question1$continent == 'Europe') %>%
  dplyr::select(continent, pf_law, pf_security, pf_movement, pf_religion, pf_assembly, pf_expression, pf_identity, ef_government, ef_legal, ef_money, ef_trade, ef_regulation)

data_Q1_Europe_HFI_long <- melt(data_Q1_Europe_HFI)

medians_EU_HFI <- data_Q1_Europe_HFI_long %>%
  group_by(variable) %>%
  summarize(median_value = median(value))

data_Q1_Europe_HFI_long <- data_Q1_Europe_HFI_long %>%
                  left_join(medians_EU_HFI, by = "variable")

#Oceania 
data_Q1_Oceania_HFI <- data_question1 %>%
  filter(data_question1$continent == 'Oceania') %>%
  dplyr::select(continent, pf_law, pf_security, pf_movement, pf_religion, pf_assembly, pf_expression, pf_identity, ef_government, ef_legal, ef_money, ef_trade, ef_regulation)

data_Q1_Oceania_HFI_long <- melt(data_Q1_Oceania_HFI)

medians_OC_HFI <- data_Q1_Oceania_HFI_long %>%
  group_by(variable) %>%
  summarize(median_value = median(value))

data_Q1_Oceania_HFI_long <- data_Q1_Oceania_HFI_long %>%
                  left_join(medians_OC_HFI, by = "variable")

# merge all medians 
medians_all_HFI <- rbind(data_Q1_Oceania_HFI_long, data_Q1_Americas_HFI_long,data_Q1_Africa_HFI_long,data_Q1_Asia_HFI_long,data_Q1_Europe_HFI_long)

medians_all_HFI$color <- ifelse(medians_all_HFI$median_value > 7.5, "lightgreen", 
                        ifelse(medians_all_HFI$median_value < 2.5, "red3", 'lightblue3'))

bandwidth_nrd_HFI <- bw.nrd(medians_all_HFI$value)

# Create the plot
ggplot(medians_all_HFI, aes(x = variable, y = value, fill = color)) +
  geom_violin(trim = FALSE, bw = bandwidth_nrd_HFI) +
  #geom_boxplot(width = 0.1, outlier.size = 1, fill = 'white') +
  scale_fill_identity() +
  labs(title = "Human Freedom Index Scores by Region", x = "Variables", y = "Score") +
  facet_grid(continent ~ ., scales = "free_y") +
  scale_y_continuous(labels = scales::label_number()) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
Here we can notice the same results as before concerning the SDG goals, except that no score has a median below 25%. Again, Europe is the continent with most of its median scores superior to 75 (lightblue color)

For space reason because of the different scales, we have decided not to make violin boxplot per continent for the remaining variables. The distribution can be seen in the general distribution seen prior to that. 

Now, let's have a closer look to the general correlation between our variables. Using our cleaned dataset, we will use a correlation heatmap to help us vizualising the informations. Given that most of our variables are not normally distributed, we will use the Spearman method to calculate the correlation.

```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE}
#### Correlations between variables Heatmap ####

Correlation_overall <-data_question1 %>% # selection of the numerical data
      dplyr::select(population:ef_regulation)

cor_matrix_sper <- # calculation of the correlation matrix
  cor(Correlation_overall, method = "spearman", use = "everything")

cor_melted <- # wide to long transformation
  melt(cor_matrix_sper)

ggplot(data = cor_melted, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name="Spearman\nCorrelation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 8, hjust = 1),
        axis.text.y = element_text(size = 8)) +
  coord_fixed() +
  labs(x = '', y = '', title = 'Correlation Matrix Heatmap')
```
By looking at our heatmap, we notice that most of our goals are strongely correlated together and that some variables amongst the Human Freedom Index scores too (strong correlation among personal freedom variables (pf), reflecting scores from the Human Freedom Index on movement, religion, assembly, and expression). This could be explained by the fact that some of these goals/scores share partially similar objectifs, which could mean that a raise in the score of one of these goals will raise positively the score of another/some other goals. In addition, we notice that goals 12 and 13 (respectively "responsible consumption & production" and "climate action") are strongely negatively correlated with most of our variables, except between themself. 

We will see more in detail the correlations between our goals and variables in the analysis part of the influence of the factors over the Sustainable Development Goals. 

In order to have an overview of the relationship between our independent variables and the SDG overall score, we make several graphs containing the Spearman correlation coefficient between the variable, the scatter plots describing the relationship between the variables, as well as the distribution of each variable. -->

```{r, echo = TRUE, message = FALSE, warning=FALSE, out.height="100%", out.width="100%"}
#### Spearman's correlation coeff ####

lower.panel <- function(x, y, ...){
   points(x, y, pch = 20, col = "darkgreen", cex = 0.2)
}
 
 panel.hist <- function(x, ...){
   usr <- par("usr"); on.exit(par(usr))
   par(usr = c(usr[1:2], 0, 1.5) )
   h <- hist(x, plot = FALSE)
   breaks <- h$breaks; nB <- length(breaks)
   y <- h$counts; y <- y/max(y)
   rect(breaks[-nB], 0, breaks[-1], y, col = "lightgreen", ...)
 }
 
 # panel.cor_stars function with stars alongside correlation coefficients
 panel.cor_stars <- function(x, y, digits = 2, prefix = "", cex.cor, ...) {
   usr <- par("usr"); on.exit(par(usr))
   par(usr = c(0, 1, 0, 1))
   r <- cor(x, y)
   p_value <- cor.test (x,y)$p.value
 
   if (p_value < 0.001){
     stars <- "***"
   } else if (p_value < 0.01) {
     stars <- "**"
   } else if (p_value < 0.05) {
     stars <- "*"
   } else {
     stars <- ""
   }
   txt <- paste0(format(c(r, 0.123456789), digits = digits)[1], " ", stars)
   if(missing(cex.cor)) cex.cor <- 0.5/strwidth(txt)
   text(0.5, 0.5, txt, cex = cex.cor)
 }
 
 # # Independent variables
 pairs(Correlation_overall[,c("overallscore", "unemployment.rate", "GDPpercapita", "MilitaryExpenditurePercentGDP", "internet_usage")], upper.panel=panel.cor_stars, diag.panel=panel.hist, lower.panel = lower.panel, main="Correlation table and distribution of various variables")
 
 # pairs(Correlation_overall[,c("overallscore", "pf_law", "pf_security", "pf_movement", "pf_religion", "pf_assembly" ,"pf_expression" ,"pf_identity", "ef_government", "ef_legal", "ef_money", "ef_trade", "ef_regulation")], upper.panel=panel.cor_stars, diag.panel=panel.hist, lower.panel = lower.panel, main="Correlation table and distribution of HFI variables")
```
Meaning of the stars: *** : p_value < 0.001; ** : p_value < 0.01; *: p_value <0.05; no star if p_value is higher.


The overall SDG achievement score is highly correlated with the percentage of people using the internet (r=.79) and GDP per capita (r=.60). The unemployement rate as well as the military expenditures in percentage of GDP per capita do not seem to play a role. However, this is only for the overall score.

The overall SDG achievement score is highly correlated with "personal freedom: law" (p=.69) and "personal freedom: identity" (p=.62). The other dimensions of personal freedom do not seem to have important influence. Regarding the distribution of the personal freedom variables, we notice that except for law, all have right-skewed distributions meaning that most of the countries have high scores.

The overall SDG achievement score is highly correlated with "economical freedom: legal" (p=.77), "economical trade: legal" (p=.67) and "economical freedom: money" (p=.6), while the other dimensions of economic freedom do not seem to have important influence. Regarding the distribution of the economic freedom variables, we notice more heterogeneous distributions and scores across the various countries than for personal freedom.

### Looking at SDGs

As we can see in the graph, most of our goals are correlated toghether. We will nown perform a PCA analysis to see how our variables are explained. You can see below the Scree plot of our PCA analysis.

```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.height="100%", out.width="100%"}
#### PCA and PCA Scree plot####

myPCA_g <- PCA(data_question1[,9:24], graph = FALSE)
fviz_eig(myPCA_g,
         addlabels = TRUE) +
  theme_minimal()
```

As we can see, Dimension 1 explain already more than 60% of the variations in our data. With Dimension 2, it goes up to around 70%. We can now plot our data with our two firsts dimensions.

```{r, echo = TRUE, message = FALSE, warning=FALSE, cache = TRUE, out.height="100%", out.width="100%"}
#### PCA Biplot ####

fviz_pca_biplot(myPCA_g,
                label="var",
                col.var="dodgerblue3",
                geom="point",
                pointsize = 0.1,
                labelsize = 5) +
  theme_minimal()
```

<<<<<<< HEAD
Concerning the SDG goals, we conclude that most of our variables are going along the 1st component, except the goals 10 and 15 that are rather uncorrelated with the dimension 1. In addition, as seen before, the goals 12 and 13 are negatively correlated to the other goals. With a eigenvalue bigger than 1 for the first two components, we conclude that there are only 2 dimensions to take into account, according to the Kaiser-Guttman's rule. Nevertheless, they are explaining less than 80% of cumulated variance.
=======
Concerning the SDG goals, we conclude that most of our variables are going along the 1st component, except the goals 10 and 15 that are rather uncorrelated with the dimension 1. In addition, as seen before, the goals 12 and 13 are negatively correlated to the other goals. With a eigenvalue bigger than 1 for the first two components, we conclude that there are only 2 dimensions to take into account, according to the Kaiser-Guttman's rule. Nevertheless, they are explaining less than 80% of cumulated variance.
