```{r echo = FALSE, message = FALSE}
#### Setup ####
source(here::here("scripts","Clean","setup.R"))
data_4 <- read.csv(here("scripts", "data", "data_question24.csv"))
threshold_heatmap <- 0.5
library(leaps)
```

## Focus on the relationships among the SDGs
**How are the different SDGs linked? (We want to see if some SDGs are linked in the fact that a high score on one implies a high score on the other, and thus if we can make groups of SDGs that are comparable in that way).**

### EDA: General visualization of the SDGs

To better analyse the relationships between the SDGs, we will first visualize the correlation between the correlation between the SDGs with the help of a heatmap. We chose to set a threshold at |`r threshold_heatmap`| to concentrate our analysis on the most linked SDGs. We initially intended to use the Pearson correlation method, but our data is, as seen in previous chapter, not normally ditributed. We tryed to normalized our data through logarithmic or square root transformation, but it was not sufficient. For that reason, we chose to use th Spearman correlation. While not being an ideal method, the Spearman correaltion does not require the data to be normally ditributed.

To do that, we select only the colums of interest and compute the correlation matrix using Spearman correlation. We then melt the matrix to be able to plot it. We then plot the heatmap using `ggplot2`.

```{r echo = TRUE, message = FALSE}
#### Preparation of the data ####

# Keeping only the columns of interest for the correlation calcluation
data_4_goals <- data_4 %>%
  dplyr::select(overallscore, goal1, goal2, goal3, goal4, goal5,
                goal6,goal7, goal8, goal9, goal10, goal11, goal12,
                goal13, goal15, goal16, goal17)
```

```{r echo = TRUE, message = FALSE}
#### Spearman Correlation ####

# Calculate Spearman correlation
spearman_corr_4 <-cor(data_4_goals,
                      method = "spearman",
                      use = "everything")

# Apply threshold and replace values below it with NA
spearman_corr_4[abs(spearman_corr_4) < threshold_heatmap] <- NA
```

```{r echo = TRUE, message = FALSE}
#### Spearman Correlation Heatmap ####

# Melting the data
melted_corr_4 <- melt(spearman_corr_4, na.rm = TRUE)

# Creation of the heatmap
ggplot(data = melted_corr_4,
       aes(x = Var1,
           y = Var2,
           fill = value)) +
    geom_tile() +
    geom_text(aes(label = sprintf("%.2f", value)),
              vjust = 0.5,
              size=2.5) + # Adding text
    scale_fill_viridis_c(name = "Spearman\nCorrelation",
                         na.value = "white",
                         direction = 1,
                         begin = 0.1,
                         end = 1) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45,
                                     hjust = 1)) +
    labs(title = "Heatmap of Spearman Correlations for Goals", 
         x = "",
         y = "")
```

The correlation can be read on the graph. The darker the color, the stronger the correlation. If there is not colors, it means that the goals correlation does not exceed our threashold of ±`r threshold_heatmap`.

It is evident that the Sustainable Development Goals (SDGs) are intricately interconnected. However, certain goals seems to be less interrelated compared to others. We can see that SDG 1 (No Poverty) and SDG 10 (Reduced Inequalities) have a weaker correlation with the rest of the goals. Similarly, Goal 15 (Life on Land) also have a weacker interconnection with the other SDGs. It is also interesting to note that some goals are negatively correlated with others. For instance, based on the Spearman correlation, goal 12 (Responsable Consumption and Production) and goal 13 (Climate Action) are negatively correlated with the others goals. This suggest that when the higher a goal other than goal 12 or 13 is, the lower the goals 12 and 13 are. Given their similar nature, it is not surprising that they are highly correlated with each other.

### Analysis: Factor analysis and Stepwise regression applied to the SDGs

At this point, we saw that the goals were mostly correlated. We now want to see if we can group them in a smaller number of factors. To do that, we will use a principal component analysis (PCA). We will first look at the scree plot to see how many factors we should keep. We will then look at the biplot to see how the goals are grouped together.

```{r echo = TRUE, message = FALSE}
#### Scree Plot ####

viridis_colors <- viridis(10, option = "D", begin = 0, end = 0.8)

# Selecting only the goals columns
goals_data <- data_4 %>%
  dplyr::select(goal1, goal2, goal3, goal4, goal5,
                goal6,goal7, goal8, goal9, goal10, goal11, goal12,
                goal13, goal15, goal16, goal17)
goals_data_scaled <- scale(goals_data) # Scaling the data
pca_result <- prcomp(goals_data_scaled) # Running PCA

# Plotting Scree plot to visualize the importance of each principal component
fviz_eig(pca_result,
         addlabels = TRUE,
         linecolor = viridis(1,
                             option = "B",
                             begin = 0.5),
         barcolor = "black",
         barfill = viridis(10,
                           option = "D",
                           begin = 0,
                           end = 0.8)) +
  theme_minimal()

eigenvalues <- pca_result$sdev^2 # getting the third eigenvalue
```

We see clearly that the first component is the most important one. Guided by the Kaiser criterion, which advises retaining only those components with eigenvalues exceeding 1, the initial three components emerge as candidates with the third components having a eigenvalue of `r eigenvalues[3]`. We now want to see how those our two first components look on in a biplot.

```{r echo = TRUE, message = FALSE}
#### Biplot ####

# Plotting Biplot to visualize the two main dimensions
fviz_pca_biplot(pca_result,
                label = "var",
                col.var = viridis(1,
                                  option = "D",
                                  begin = 0.5),
                geom = "point",
                pointsize = 0.1,
                labelsize = 4) +
  theme_minimal()
```

The biplot offers insightful visualization, clearly illustrating the relationship between the various goals and the first two components. Notably, Dimension 2 exhibits a strong correlation with Goals 10 (Reduced inequalities) and 15 (Life on Land), whereas the remaining goals show a moderate to high correlation with Dimension 1. Furthermore, an interesting pattern emerges, revealing three distinct groups of variables, each playing a unique role. One group comprises Goals 12 (Responsible Consumption and Production) and 13 (Climate Action), another encompasses Goals 10 (Reduced inequalities) and 15 (Life on Land), and the third group includes all other variables. This categorization aids in understanding the distinct influences and interactions among the goals.

Grouping Goal 12 (Responsible Consumption and Production) and Goal 13 (Climate Action) together is logical, as both pertain to environmental issues. It is, however, interesting to note the pairing of Goal 10 (Reduced Inequalities) with Goal 15 (Life on Land). This could be explained by the fact that Goal 10 (Reduced inequalities) is related to the reduction of inequalities within and among countries, while Goal 15 (Life on Land) is related to the protection, restoration and promotion of sustainable use of terrestrial ecosystems, sustainable manage forests, combat desertification, and halt and reverse land degradation and halt biodiversity loss. Therefore, it is possible that the respondents who are more concerned about the reduction of inequalities are also more concerned about the protection of the environment. But this is a stretched.

We also performed a stepwise regression to see more precisely how the goals are correlated with each other. We use the forward selection with the `leaps` package.

```{r echo = TRUE, message = FALSE, warning=FALSE}
#### Preparation of the data ####

goals_data <- data_4 %>%
  dplyr::select(overallscore, goal1, goal2, goal3, goal4, goal5,
                goal6,goal7, goal8, goal9, goal10, goal11, goal12,
                goal13, goal15, goal16, goal17)
```

```{r echo = TRUE, message = FALSE, warning=FALSE}
#### Stepwise regression ####

leaps_o <- regsubsets(overallscore ~ goal1 + goal2 + goal3 + goal4 + goal5 +
                        goal6 + goal7 + goal8 + goal9 + goal10 + goal11 +
                        goal12 + goal13 + goal15 + goal16 + goal17,
                      data = goals_data, nbest=1, method = "forward")
leaps_1 <- regsubsets(goal1 ~ goal2 + goal3 + goal4 + goal5 + goal6 + goal7 +
                        goal8 + goal9 + goal10 + goal11 + goal12 + goal13 +
                        goal15 + goal16 + goal17, data = goals_data, nbest=1,
                      method = "forward")
leaps_2 <- regsubsets(goal2 ~ goal1 + goal3 + goal4 + goal5 + goal6 + goal7 +
                        goal8 + goal9 + goal10 + goal11 + goal12 + goal13 +
                        goal15 + goal16 + goal17, data = goals_data, nbest=1,
                      method = "forward")
leaps_3 <- regsubsets(goal3 ~ goal1 + goal2 + goal4 + goal5 + goal6 + goal7 +
                        goal8 + goal9 + goal10 + goal11 + goal12 + goal13 +
                        goal15 + goal16 + goal17, data = goals_data, nbest=1,
                      method = "forward")
leaps_4 <- regsubsets(goal4 ~ goal1 + goal2 + goal3 + goal5 + goal6 + goal7 +
                        goal8 + goal9 + goal10 + goal11 + goal12 + goal13 +
                        goal15 + goal16 + goal17, data = goals_data, nbest=1,
                      method = "forward")
leaps_5 <- regsubsets(goal5 ~ goal1 + goal2 + goal3 + goal4 + goal6 + goal7 +
                        goal8 + goal9 + goal10 + goal11 + goal12 + goal13 +
                        goal15 + goal16 + goal17, data = goals_data, nbest=1,
                      method = "forward")
leaps_6 <- regsubsets(goal6 ~ goal1 + goal2 + goal3 + goal4 + goal5 + goal7 +
                        goal8 + goal9 + goal10 + goal11 + goal12 + goal13 +
                        goal15 + goal16 + goal17, data = goals_data, nbest=1,
                      method = "forward")
leaps_7 <- regsubsets(goal7 ~ goal1 + goal2 + goal3 + goal4 + goal5 + goal6 +
                        goal8 + goal9 + goal10 + goal11 + goal12 + goal13 +
                        goal15 + goal16 + goal17, data = goals_data, nbest=1,
                      method = "forward")
leaps_8 <- regsubsets(goal8 ~ goal1 + goal2 + goal3 + goal4 + goal5 + goal6 +
                        goal7 + goal9 + goal10 + goal11 + goal12 + goal13 +
                        goal15 + goal16 + goal17, data = goals_data, nbest=1,
                      method = "forward")
leaps_9 <- regsubsets(goal9 ~ goal1 + goal2 + goal3 + goal4 + goal5 + goal6 +
                        goal7 + goal8 + goal10 + goal11 + goal12 + goal13 +
                        goal15 + goal16 + goal17, data = goals_data, nbest=1,
                      method = "forward")
leaps_10 <- regsubsets(goal10 ~ goal1 + goal2 + goal3 + goal4 + goal5 + goal6 +
                         goal7 + goal8 + goal9 + goal11 + goal12 + goal13 +
                         goal15 + goal16 + goal17, data = goals_data, nbest=1,
                       method = "forward")
leaps_11 <- regsubsets(goal11 ~ goal1 + goal2 + goal3 + goal4 + goal5 + goal6 +
                         goal7 + goal8 + goal9 + goal10 + goal12 + goal13 +
                         goal15 + goal16 + goal17, data = goals_data, nbest=1,
                       method = "forward")
leaps_12 <- regsubsets(goal12 ~ goal1 + goal2 + goal3 + goal4 + goal5 + goal6 +
                         goal7 + goal8 + goal9 + goal10 + goal11 + goal13 +
                         goal15 + goal16 + goal17, data = goals_data, nbest=1,
                       method = "forward")
leaps_13 <- regsubsets(goal13 ~ goal1 + goal2 + goal3 + goal4 + goal5 + goal6 +
                         goal7 + goal8 + goal9 + goal10 + goal11 + goal12 +
                         goal15 + goal16 + goal17, data = goals_data, nbest=1,
                       method = "forward")
leaps_15 <- regsubsets(goal15 ~ goal1 + goal2 + goal3 + goal4 + goal5 + goal6 +
                         goal7 + goal8 + goal9 + goal10 + goal11 + goal12 +
                         goal13 + goal16 + goal17, data = goals_data, nbest=1,
                       method = "forward")
leaps_16 <- regsubsets(goal16 ~ goal1 + goal2 + goal3 + goal4 + goal5 + goal6 +
                         goal7 + goal8 + goal9 + goal10 + goal11 + goal12 +
                         goal13 + goal15 + goal17, data = goals_data, nbest=1,
                       method = "forward")
leaps_17 <- regsubsets(goal17 ~ goal1 + goal2 + goal3 + goal4 + goal5 + goal6 +
                         goal7 + goal8 + goal9 + goal10 + goal11 + goal12 +
                         goal13 + goal15 + goal16, data = goals_data, nbest=1,
                       method = "forward")
```

```{r echo = TRUE, message = FALSE, warning=FALSE}
#### Regression found with stepwise regression ####

mod_o <- lm(
  overallscore ~ goal2 + goal3 + goal4 + goal6 + goal7 + goal10 + goal15 + goal17,
  data = goals_data)
mod_1 <- lm(
  goal1 ~ goal3 + goal4 + goal5 + goal6 + goal7 + goal9 + goal13 + goal17,
  data = goals_data)
mod_2 <- lm(
  goal2 ~ goal4 + goal5 + goal6 + goal8 + goal9 + goal12 + goal16 + goal17,
  data = goals_data)
mod_3 <- lm(
  goal3 ~ goal1 + goal4 + goal7 + goal8 + goal9 + goal11 + goal15 + goal16,
  data = goals_data)
mod_4 <- lm(
  goal4 ~ goal1 + goal2 + goal3 + goal5 + goal7 + goal11 + goal16 + goal17,
  data = goals_data)
mod_5 <- lm(
  goal5 ~ goal1 + goal4 + goal6 + goal9 + goal10 + goal11 + goal15 + goal17,
  data = goals_data)
mod_6 <- lm(
  goal6 ~ goal1 + goal2 + goal3 + goal5 + goal8 + goal9 + goal11 + goal15,
  data = goals_data)
mod_7 <- lm(
  goal7 ~ goal1 + goal3 + goal4 + goal5 + goal6 + goal8 + goal11 + goal13,
  data = goals_data)
mod_8 <- lm(
  goal8 ~ goal2 + goal5 + goal6 + goal9 + goal10 + goal12 + goal15 + goal17,
  data = goals_data)
mod_9 <- lm(
  goal9 ~ goal1 + goal2 + goal3 + goal8 + goal10 + goal12 + goal13 + goal17,
  data = goals_data)
mod_10 <- lm(
  goal10 ~ goal1 + goal5 + goal9 + goal11 + goal13 + goal15 + goal16 + goal17,
  data = goals_data)
mod_11 <- lm(
  goal11 ~ goal3 + goal4 + goal5 + goal6 + goal7 + goal10 + goal15 + goal16,
  data = goals_data)
mod_12 <- lm(
  goal12 ~ goal2 + goal7 + goal8 + goal9 + goal13 + goal15 + goal16 + goal17,
  data = goals_data)
mod_13 <- lm(
  goal13 ~ goal1 + goal5 + goal7 + goal9 + goal10 + goal12 + goal15 + goal16,
  data = goals_data)
mod_15 <- lm(
  goal15 ~ goal3 + goal4 + goal5 + goal6 + goal10 + goal11 + goal12 + goal13,
  data = goals_data)
mod_16 <- lm(
  goal16 ~ goal2 + goal3 + goal4 + goal10 + goal11 + goal12 + goal13 + goal17,
  data = goals_data)
mod_17 <- lm(
  goal17 ~ goal1 + goal5 + goal8 + goal9 + goal10 + goal11 + goal12 + goal16,
  data = goals_data)
```

Below, we create a graph to present our different model that the stepwise regression chose as best regression for each of our goals. As expected, the model that explain the variable overallscore is explained by the other goals at a similar coefficient for each of our goals. This make sense since the overallscore has been directly calculated with the 17 scores. 

```{r echo = TRUE, message = FALSE, warning=FALSE}
library('broom')
# Create a dataframe of tidy models
model_list <-
  list(mod_o, mod_1, mod_2, mod_3, mod_4, mod_5,
                   mod_6, mod_7, mod_8, mod_9, mod_10, mod_11,
                   mod_12, mod_13, mod_15, mod_16, mod_17)

models_tidy <-
  lapply(model_list, tidy)

names(models_tidy) <-
  c("Overallscore ~others", "Goal1 ~ others", "Goal2 ~ others",
    "Goal3 ~ others", "Goal4 ~ others", "Goal5 ~ others", "Goal6 ~ others",
    "Goal7 ~ others", "Goal8 ~ others", "Goal9 ~ others", "Goal10 ~ others",
    "Goal11 ~ others", "Goal12 ~ others", "Goal13 ~ others", "Goal15 ~ others",
    "Goal16 ~ others", "Goal17 ~ others")

# Combine into a single dataframe
df_tidy <-
  do.call(rbind,
          lapply(names(models_tidy),
                 function(x) cbind(models_tidy[[x]], Model=x)))

# Assuming 'p.value' is the column name for p-values in your dataframe
significance_level <- 0.05

# Filter for significant p-values
df_tidy_significant <-
  df_tidy[df_tidy$p.value < significance_level, ]

# Plot
ggplot(df_tidy_significant,
       aes(x = Model,
           y = estimate,
           color = term)) +
  geom_point() +
  geom_errorbar(aes(ymin = estimate - std.error,
                    ymax = estimate + std.error),
                width = 0.2) +
  ylim(-1, 1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right", # You can change the position if needed
        legend.text = element_text(size = 5), # Adjust text size
        legend.title = element_text(size = 7), # Adjust title size
        legend.key.size = unit(0.3, "cm")) +  # Adjust key size) +
  labs(title = "Coefficient Plot of Regression Models",
       x = "Models",
       y = "Estimates")
```

```{r echo = TRUE, message = FALSE, warning=FALSE}
plot(mod_o, which = 1)
plot(mod_1, which = 1)
plot(mod_2, which = 1)
plot(mod_3, which = 1)
plot(mod_4, which = 1)
plot(mod_5, which = 1)
plot(mod_6, which = 1)
plot(mod_7, which = 1)
plot(mod_8, which = 1)
plot(mod_9, which = 1)
plot(mod_10, which = 1)
plot(mod_11, which = 1)
plot(mod_12, which = 1)
plot(mod_13, which = 1)
plot(mod_15, which = 1)
plot(mod_16, which = 1)
plot(mod_17, which = 1)
```

